---
title: "snow_albedo_processing"
author: "Erika Lee"
date: "2024-05-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#packages
library(tidyverse)
library(lubridate)
library(plotly)
library(ggplot2)
library(stringr)
library(readxl)
library(rstatix)
library(plotly)
library(patchwork)
library(ggpubr)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

```{r}
#read in weather data for burned stations
trans_burned_wx_15min_r = read_csv('nsf/trans_burned/trans_burned_wx_15min_r.csv') %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")), datetime = force_tz(datetime, tz = 'MST'))

pers_burned_wx_hourly_r = read_csv('nsf/pers_burned/pers_burned_wx_hourly_r.csv') %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")), datetime = force_tz(datetime, tz = 'MST'))
```

```{r}
#read in persistent burned data
pers_ub_albedo_all <- read_excel("nsf/albedo/Pers_ub_albedo.xlsx") %>%
  mutate(date = as_date(Date), time = format(Time, "%H:%M:%S")) %>%
  mutate(datetime = as.POSIXct(paste(date, time))) %>%
  mutate(datetime = force_tz(datetime, tzone = "MST")) %>%
  select(-c(Date, Time)) %>%
  select(date, time, everything())

#read in persistent data
pers_albedo_all <- read_excel("nsf/albedo/Pers_b_albedo.xlsx") %>%
mutate(date = as_date(Date), time = format(Time, "%H:%M:%S")) %>%
  mutate(datetime = as.POSIXct(paste(date, time))) %>%
  mutate(datetime = force_tz(datetime, tzone = "MST")) %>%
  select(-c(Date, Time)) %>%
  select(date, time, everything())

#read in persistent data - this is V2, where I manipulated the 5/31 data!
pers_albedo_all_V2 <- read_excel("nsf/albedo/Pers_b_albedo_V2.xlsx") %>%
  # Convert Date to the correct date format
  mutate(date = as_date(Date), time = format(Time, "%H:%M:%S")) %>%
  mutate(datetime = as.POSIXct(paste(date, time))) %>%
  mutate(datetime = force_tz(datetime, tzone = "MST")) %>%
  select(-c(Date, Time)) %>%
  select(date, time, everything())

#selecting for May 13 albedo readings
pers_albedo_may13 <- pers_albedo_all %>%
  filter(date == '2024-05-13')

#selecting for May 31 albedo readings
pers_albedo_may31 <- pers_albedo_all_V2 %>%
  filter(date == '2024-05-31')

pers_mean_albedo_by_date <- pers_albedo_all %>%
  group_by(date) %>%
  summarize(mean_Albedo = mean(Albedo, na.rm = TRUE))

# Join the mean albedo data back to the original dataframe
pers_albedo_all <- pers_albedo_all %>%
  left_join(pers_mean_albedo_by_date, by = "date") %>%
  select(datetime, date, time, everything())
  

# Read the Excel file
trans_albedo_all <- read_excel("nsf/albedo/Trans_b_albedo.xlsx") %>%
  mutate(date = as_date(Date), time = format(Time, "%H:%M:%S")) %>%
  mutate(datetime = as.POSIXct(paste(date, time))) %>%
  mutate(datetime = force_tz(datetime, tzone = "MST")) %>%
  select(-c(Date, Time)) %>%
  select(date, time, everything())

#selecting for April albedo readings
trans_albedo_april <- trans_albedo_all %>%
  filter(date == '2024-04-12')

trans_albedo_feb <- trans_albedo_all %>%
  filter(date == '2024-02-19')


trans_mean_albedo_by_date <- trans_albedo_all %>%
  group_by(date) %>%
  summarize(mean_Albedo = mean(Albedo, na.rm = TRUE))

# Join the mean albedo data back to the original dataframe
trans_albedo_all <- trans_albedo_all %>%
  left_join(trans_mean_albedo_by_date, by = "date") %>%
  select(datetime, date, time, everything())

#creating a combined dataset with SWin and SWout together
trans_SW_combined <- trans_albedo_all %>%
  pivot_longer(cols = starts_with("SW"), 
               names_to = "SW",
               values_to = "value") %>%
  #filter out only those that say sun
  filter(Sun_shadow == "sun")
```

## Newer Dataframes & Plotting (2024.8.28)

Creating cleaned up, averaged by month, dataframes for new plots - as of 2024.08.28

```{r}
#transitional zone
avg_trans_albedo_lowmort <- trans_albedo_all %>%
  filter(Site == "trans_b_lp_lowmort") %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarize(mean_swin = mean(SWin_wperm2), mean_swout = mean(SWout_wperm2), mean_albedo = mean(mean_Albedo.x)) %>%
  #adding a date column so I can plot with weather station data
  mutate(date = as.Date(paste0("2024-", month, "-15"), format = "%Y-%m-%d")) %>%
  mutate(date = as.Date(date))

avg_trans_albedo_highmort <- trans_albedo_all %>%
  filter(Site == "trans_b_lp_highmort") %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarize(mean_swin = mean(SWin_wperm2), mean_swout = mean(SWout_wperm2), mean_albedo = mean(mean_Albedo.x)) %>%
  mutate(date = as.Date(paste0("2024-", month, "-15"), format = "%Y-%m-%d")) %>%
  mutate(date = as.Date(date))
```

\*\*I don't have persistent data for any months except May, so these dataframes are not that helpful. Once I have better data then I can compare them better?

```{r}
#persistent zone
avg_pers_albedo_lowmort <- pers_albedo_all %>%
  filter(Site == "pers_b_lp_lowmort") %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarize(mean_swin = mean(SWin_wperm2), mean_swout = mean(SWout_wperm2), mean_albedo = mean(mean_Albedo.x)) %>%
  #adding a mid-month date column so I can plot with the weather station data
  mutate(date = "2024-05-15") %>%
  mutate(date = as.Date(date))

avg_pers_albedo_highmort <- pers_albedo_all %>%
  filter(Site == "pers_b_lp_highmort") %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarize(mean_swin = mean(SWin_wperm2), mean_swout = mean(SWout_wperm2), mean_albedo = mean(mean_Albedo.x)) %>%
  mutate(date = "2024-05-15") %>%
  mutate(date = as.Date(date))
```

Weather Station Dataframes

```{r}
#filtering for the months of feb-may
filtered_daily_persb_wx <- pers_burned_wx_hourly_r %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00") & datetime <= as.POSIXct("2024-06-05 00:00:00")) %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date) %>%
  summarize(avg_swin_wperm2 = mean(SWin_Avg), avg_swout_wperm2 = mean(SWout_Avg), mean_sw_albedo = mean(SWalbedo_Avg), mean_sdepth_m = mean(SnowDepth_m)) %>%
  mutate(mean_sdepth_cm = mean_sdepth_m*100)

filtered_daily_transb_wx <- trans_burned_wx_15min_r %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00") & datetime <= as.POSIXct("2024-06-05 00:00:00")) %>%
  drop_na() %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date) %>%
  summarize(avg_swin_wperm2 = mean(SWin_Avg), avg_swout_wperm2 = mean(SWout_Avg), mean_sw_albedo = mean(SWalbedo_Avg), mean_sdepth_m = mean(SnowDepth_m)) %>%
  mutate(mean_sdepth_cm = mean_sdepth_m*100)
```

Plottting Wx Station Data and monthly albedo plot readings

```{r}
#setting constants for plots
color_values <- c(
  "avg_swin_wperm2" = "#D95F02",
  "avg_swout_wperm2" = "#E6AB02",
  "mean_sw_albedo" = "#66A61E",
  "mean_sdepth_cm" = "#7570B3", 
  "plot_albedo_highmort" = "red", 
  "plot_albedo_lowmort" = "black")

# Define line types
line_types <- c(
  "Weather Station" = "solid",
  "Tree Temperature Sensors" = "dashed"
)

#changing labels of values in plots for the legend
color_labels <- c(
  "avg_swin_wperm2" = "Wx Station Shortwave In",
  "avg_swout_wperm2" = "Wx Station Shortwave Out",
  "mean_sw_albedo" = "Wx Station Shortwave Albedo", 
  "plot_albedo_highmort" = "High Mortality Plot Albedo",
  "plot_albedo_lowmort" = "Low Mortality Plot Albedo"
)
```

```{r}
pers_sw_wx_plot_data <- ggplot() +  
  geom_line(data = filtered_daily_persb_wx, aes(x = date, y = avg_swin_wperm2, color = "avg_swin_wperm2")) +
  geom_line(data = filtered_daily_persb_wx, aes(x = date, y = avg_swout_wperm2, color = "avg_swout_wperm2")) +
  labs(
    title = "Persistent Burned",
    x = "Date",
    y = "Radiation components (W/m2)",
    color = "Variable"
    ) +
  theme_bw() +  # Use a minimal theme for a clean look
  scale_color_manual(values = color_values, labels = color_labels)

pers_sw_wx_plot_data

pers_albedo_plot_data <- ggplot() +  
 geom_line(data = filtered_daily_persb_wx, aes(x = date, y = mean_sw_albedo, color = "mean_sw_albedo")) +
  # Add points for avg_swout_wperm2
  geom_point(data = avg_pers_albedo_highmort, 
             aes(x = date, y = mean_albedo, color = "plot_albedo_highmort"), size = 2) +
  geom_point(data = avg_pers_albedo_lowmort, 
             aes(x = date, y = mean_albedo, color = "plot_albedo_lowmort"), size = 1) +
  labs(
    x = "Date",
    y = "Albedo (unitless)",
    color = "Variable"
    ) +
  theme_bw() +  # Use a minimal theme for a clean look
  scale_color_manual(values = color_values, labels = color_labels)

trans_albedo_plot_data
```

```{r}
#swin/out weather station data - to be stacked ontop of albedo plot because y-values are very different
trans_sw_wx_plot_data <- ggplot() +  
  geom_line(data = filtered_daily_transb_wx, aes(x = date, y = avg_swin_wperm2, color = "avg_swin_wperm2")) +
  geom_line(data = filtered_daily_transb_wx, aes(x = date, y = avg_swout_wperm2, color = "avg_swout_wperm2")) +
  labs(
    title = "Transitional Burned",
    x = "Date",
    y = "Radiation components (W/m2)",
    color = "Variable"
    ) +
  theme_bw() +  # Use a minimal theme for a clean look
  scale_color_manual(values = color_values, labels = color_labels)

trans_sw_wx_plot_data

trans_albedo_plot_data <- ggplot() +  
 geom_line(data = filtered_daily_transb_wx, aes(x = date, y = mean_sw_albedo, color = "mean_sw_albedo")) +
  # Add points for avg_swout_wperm2
  geom_point(data = avg_trans_albedo_highmort, 
             aes(x = date, y = mean_albedo, color = "plot_albedo_highmort"), size = 2) +
  geom_point(data = avg_trans_albedo_lowmort, 
             aes(x = date, y = mean_albedo, color = "plot_albedo_lowmort"), size = 1) +
  labs(
    x = "Date",
    y = "Albedo (unitless)",
    color = "Variable"
    ) +
  theme_bw() +  # Use a minimal theme for a clean look
  scale_color_manual(values = color_values, labels = color_labels)

trans_albedo_plot_data
```

```{r}
#arranging like plots next to each other, then combining for a final plot

comb_sw_wx_plots <- ggarrange(
  pers_sw_wx_plot_data, 
  trans_sw_wx_plot_data, 
  nrow = 1, 
  common.legend = TRUE, 
  legend = "right"
)

comb_sw_wx_plots

comb_albedo_plots <- ggarrange(pers_albedo_plot_data, trans_albedo_plot_data, nrow = 1, 
  common.legend = TRUE, 
  legend = "right"
)

comb_albedo_plots
```

```{r}
#final combined plot
final_comb_swalbedo_plot <- ggarrange(comb_sw_wx_plots, comb_albedo_plots, nrow = 2, common.legend = TRUE, legend = "right")

final_comb_swalbedo_plot
```

```{r}
#saving combined plot
ggsave(filename = "final_comb_swalbedo_plot.png", plot = final_comb_swalbedo_plot, width = 15, height = 10, dpi = 300)
```

## Older Dataframes, Wrangling & Plotting

```{r}
#plot data
trans_albedo <- ggplot(data = trans_SW_combined) +
  geom_point(aes(x = Record_location, y = value, color = SW)) +
  ggtitle("transitional")

# Convert the ggplot object to a plotly object
trans_albedo_plotly <- ggplotly(trans_albedo)

# Print the plotly object to display the interactive plot
trans_albedo_plotly
```

Snow depth and albedo plots

```{r}
#transitional snow depth
transb_snowd_plot <- ggplot() +
  geom_line(data = trans_burned_wx_15min_r, aes(x = datetime, y = SnowDepth_m), color = "black") +
  labs(
    title = "Transitional Snow Depth Over Time",
    x = "Date",
    y = "Snow Depth (m)",
  ) +
  theme_minimal()

transb_snowd_plot

#persistent snow depth
##* not sure if this snow depth column is the correct data to be using!
persb_snowd_plot <- ggplot() +
  geom_line(data = pers_burned_wx_15min_r, aes(x = datetime, y = DT_Avg), color = "black") +
  labs(
    title = "Persistent Snow Depth Over Time",
    x = "Date",
    y = "Snow Depth (m)",
  ) +
  theme_minimal()

persb_snowd_plot
```

Plotting albedo by transect point and date

\*\*these plots look good!

```{r}
#transitional
##april plot - both low and high mort
trans_albedo_plot_april <- ggplot() +
  geom_col(data = trans_albedo_april, aes(x = Record_location, y = Albedo, fill = Site), alpha = 0.7, width = 0.6, position = "dodge") +
  labs(
    title = "Transitional Burned Albedo: 2024-04-12",
    x = "Transect point (1 m apart)",
    y = "Albedo (unitless)"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(trans_albedo_all$Record_location), max(trans_albedo_all$Record_location), by = 1)) +
  scale_fill_manual(values = c("trans_b_lp_highmort" = "red", "trans_b_lp_lowmort" = "blue"))


trans_albedo_plot_april

#february plot
trans_albedo_plot_feb <- ggplot() +
  geom_col(data = trans_albedo_feb, aes(x = Record_location, y = Albedo, fill = Site), alpha = 0.7, width = 0.6, position = "dodge") +
  labs(
    title = "Transitional Burned Albedo: 2024-02-19",
    x = "Transect point (1 m apart)",
    y = "Albedo (unitless)"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(trans_albedo_all$Record_location), max(trans_albedo_all$Record_location), by = 1)) +
  scale_fill_manual(values = c("trans_b_lp_highmort" = "red", "trans_b_lp_lowmort" = "blue"))


trans_albedo_plot_feb

#combining plots
trans_b_albedo_plots_all<- trans_albedo_plot_feb + trans_albedo_plot_april + plot_layout(nrow = 2)
```

```{r}
#transitional SWin/out plot
##february
trans_albedo_SWinout_feb <- ggplot() +
  # SWin
  geom_line(data = trans_albedo_feb, aes(x = Record_location, y = SWin_wperm2, color = Site, group = Site, linetype = "SWin"), alpha = 0.7, linewidth = 0.8) +
  # SWout
  geom_line(data = trans_albedo_feb, aes(x = Record_location, y = SWout_wperm2, color = Site, group = Site, linetype = "SWout"), alpha = 0.5, linewidth = 0.8) +
  labs(
    x = "Transect point (1 m apart)",
    y = "SWin/out (W/m2)"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(trans_albedo_all$Record_location), max(trans_albedo_all$Record_location), by = 1)) +
  scale_color_manual(values = c("trans_b_lp_highmort" = "red", "trans_b_lp_lowmort" = "blue")) +
  scale_linetype_manual(values = c("SWin" = "solid", "SWout" = "dashed"))

trans_albedo_SWinout_feb

#april
trans_albedo_SWinout_april <- ggplot() +
  # SWin
  geom_line(data = trans_albedo_april, aes(x = Record_location, y = SWin_wperm2, color = Site, group = Site, linetype = "SWin"), alpha = 0.7, linewidth = 0.8) +
  # SWout
  geom_line(data = trans_albedo_april, aes(x = Record_location, y = SWout_wperm2, color = Site, group = Site, linetype = "SWout"), alpha = 0.5, linewidth = 0.8) +
  labs(
    x = "Transect point (1 m apart)",
    y = "SWin/out (W/m2)"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(trans_albedo_all$Record_location), max(trans_albedo_all$Record_location), by = 1)) +
  scale_color_manual(values = c("trans_b_lp_highmort" = "red", "trans_b_lp_lowmort" = "blue")) +
  scale_linetype_manual(values = c("SWin" = "solid", "SWout" = "dashed"))

trans_albedo_SWinout_april
```

```{r}
#combining february albedo and SWin/out plots
transb_albedo_allplots_feb <- trans_albedo_plot_feb + trans_albedo_SWinout_feb + plot_layout(nrow = 2)

transb_albedo_allplots_feb

#combining april albedo and SWin/out plots
transb_albedo_allplots_april <- trans_albedo_plot_april + trans_albedo_SWinout_april + plot_layout(nrow = 2)

transb_albedo_allplots_april

#combinging just february and april albedo plots
trans_b_albedo_plots_all<- trans_albedo_plot_feb + trans_albedo_plot_april + plot_layout(nrow = 2)

#saving february all plots
ggsave("transb_albedo_allplots_feb.png", plot = transb_albedo_allplots_feb, width = 8, height = 6, dpi = 300, bg = "white")

#saving april all plots
ggsave("transb_albedo_allplots_april.png", plot = transb_albedo_allplots_april, width = 8, height = 6, dpi = 300, bg = "white")

# Save the transitional combined plot as a PNG file to working directory
ggsave("trans_b_albedo_plots_all.png", plot = trans_b_albedo_plots_all, width = 8, height = 6, dpi = 300, bg = "white")
```

```{r}
#persistent - May 13th Date

pers_albedo_plot_may13 <- ggplot() +
  geom_col(data = pers_albedo_may13, aes(x = Record_location, y = Albedo, fill = Site), alpha = 0.7, width = 0.6, position = "dodge") +
  labs(
    title = "Persistent Burned Albedo: 2024-05-13",
    x = "Transect point (1 m apart)",
    y = "Albedo (unitless)"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(trans_albedo_all$Record_location), max(trans_albedo_all$Record_location), by = 1)) +
  scale_fill_manual(values = c("pers_b_lp_highmort" = "red", "pers_b_lp_lowmort" = "blue"))

pers_albedo_plot_may13

#pers_albedo_plot_may31
# Combine the datasets to use for ggplot
pers_combined_albedo_may31 <- bind_rows(
  pers_albedo_may31 %>% mutate(DataType = "Burned"),
  pers_ub_albedo_all %>% mutate(DataType = "Unburned")
)

pers_albedo_plot_may31 <- ggplot(pers_combined_albedo_may31, aes(x = Record_location, y = Albedo, fill = Site, group = interaction(Record_location, DataType))) +
  geom_col(alpha = 0.7, width = 0.4, position = position_dodge(width = 0.5)) +
  labs(
    title = "Persistent Burned & Unburned Albedo: 2024-05-31",
    x = "Transect point (1 m apart)",
    y = "Albedo (unitless)"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(trans_albedo_all$Record_location), max(trans_albedo_all$Record_location), by = 1)) +
  scale_fill_manual(values = c("pers_b_lp_highmort" = "red", "pers_ub" = "blue"))

pers_albedo_plot_may31
```

```{r}
#persistent SWin/out plot
##May 13th
pers_albedo_SWinout_may13 <- ggplot() +
  # SWin
  geom_line(data = pers_albedo_may13, aes(x = Record_location, y = SWin_wperm2, color = Site, group = Site, linetype = "SWin"), alpha = 0.7, linewidth = 0.8) +
  # SWout
  geom_line(data = pers_albedo_may13, aes(x = Record_location, y = SWout_wperm2, color = Site, group = Site, linetype = "SWout"), alpha = 0.5, linewidth = 0.8) +
  labs(
    x = "Transect point (1 m apart)",
    y = "SWin/out (W/m2)"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(trans_albedo_all$Record_location), max(trans_albedo_all$Record_location), by = 1)) +
  scale_color_manual(values = c("pers_b_lp_highmort" = "red", "pers_b_lp_lowmort" = "blue")) +
  scale_linetype_manual(values = c("SWin" = "solid", "SWout" = "dashed"))

pers_albedo_SWinout_may13

#May31
pers_albedo_SWinout_may31 <- ggplot() +
  # SWin
  geom_line(data = pers_albedo_may31, aes(x = Record_location, y = SWin_wperm2, color = Site, group = Site, linetype = "SWin"), alpha = 0.7, linewidth = 0.8) +
  # SWout
  geom_line(data = pers_albedo_may31, aes(x = Record_location, y = SWout_wperm2, color = Site, group = Site, linetype = "SWout"), alpha = 0.5, linewidth = 0.8) +
  geom_line(data = pers_ub_albedo_all, aes(x = Record_location, y = SWin_wperm2, color = Site, group = Site, linetype = "SWin"), alpha = 0.7, linewidth = 0.8) +
  geom_line(data = pers_ub_albedo_all, aes(x = Record_location, y = SWout_wperm2, color = Site, group = Site, linetype = "SWout"), alpha = 0.5, linewidth = 0.8) +
  labs(
    x = "Transect point (1 m apart)",
    y = "SWin/out (W/m2)"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(trans_albedo_all$Record_location), max(trans_albedo_all$Record_location), by = 1)) +
  scale_color_manual(values = c("pers_b_lp_highmort" = "red", "pers_ub" = "blue")) +
  scale_linetype_manual(values = c("SWin" = "solid", "SWout" = "dashed"))

pers_albedo_SWinout_may31
```

```{r}
#combining May 13th albedo and SWin/out plots
persb_albedo_allplots_may13 <- pers_albedo_plot_may13 + pers_albedo_SWinout_may13 + plot_layout(nrow = 2)

persb_albedo_allplots_may13

#combining May 31st albedo and SWin/out plots
pers_albedo_allplots_may31 <- pers_albedo_plot_may31 + pers_albedo_SWinout_may31 + plot_layout(nrow = 2)

pers_albedo_allplots_may31

#saving May 13th all plots
ggsave("persb_albedo_allplots_may13.png", plot = persb_albedo_allplots_may13, width = 8, height = 6, dpi = 300, bg = "white")

#saving May 31st all plots
ggsave("pers_albedo_allplots_may31.png", plot = pers_albedo_allplots_may31, width = 8, height = 6, dpi = 300, bg = "white")
```
